---
title: "Lizard Learning Plot Manuscript"
author: "Caleigh Charlebois"
date: "2022-09-15"
output: pdf_document
latex_engine: "xelatex"
geometry: margin=1in
csl: "../references/biology-letters.csl"
bibliography: "../references/lizard_learning_plot.bib"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Use groundhog to ensure package compatibility by loading most recent
## versions of packages available on a certain date.

# Check whether groundhog is installed and install it if it isn't.
my_packages <- c("groundhog")
not_installed <- my_packages[!(my_packages %in% installed.packages()[ , "Package"])]
if(length(not_installed)) install.packages(not_installed) 
rm(not_installed, my_packages)

# Load groundhog and use it to install other packages. You may change the date
# in the "date_for_packages" if you want to check results using another set of
# plugin versions.
date_for_packages <- "2022-09-11"
library(groundhog)
groundhog.library(c("ggplot2",
                    "readr",
                    "cowplot",
                    "dplyr"),
                  date_for_packages)
rm(date_for_packages, r.need.full)
```

**Title:** Lizard Learning Plot

Caleigh Charlebois


# Abstract

# Introduction

# Methods

# Results
```{r}
lizards <- read_csv("../data/data_T1_final_downloaded_2022-09-14.csv")

ggplot( data = lizards, mapping = aes()) +
  geom_histogram(aes(x=LT)) +
  facet_grid(rows = vars(Age))
```
This was my initial plan -- Making a plot of the distribution of number of trials it took for the lizards to perform the task correctly, faceted by age. However, without polishing the plot and fixing the labels, I can see it won't be very exciting and isn't the best way to visualize this data.I didn't realize when I was making my plan that the  lizards would all learn the task in either 0 or 1 trial. 

What else can I make a plot of which will be more interesting?

```{r}
lizards <- read_csv("../data/data_T1_final_downloaded_2022-09-14.csv")

ggplot(data = lizards, mapping = aes()) +
  geom_histogram(aes(x=Lat_min1)) +
  facet_grid(rows = vars(Age),
             cols = vars(Treat)) +
  labs(x="Time until correctly performed task",
       y="Lizard count") +
  theme_cowplot()
```
I find this idea pretty interesting, but I notice there are many data points at "60." I think this is because lizards who never completed the task correctly were recorded as taking 60 seconds, the maximum allotted time. Next, I will try creating a plot that removes lizards which never succeeded in performing the task by filtering the data.

```{r}
lizards <- read_csv("../data/data_T1_final_downloaded_2022-09-14.csv")

filter(lizards, correct.incorrect==1) %>%

ggplot(mapping = aes()) +
  geom_histogram(aes(x=Lat_min1)) +
  facet_grid(rows = vars(Age),
             cols = vars(Treat)) +
  theme_cowplot()
```

# Discussion

# References

# Tables

# Figure Captions

# Figures

# Appendices
