---
title: "Lizard Learning Plot Manuscript"
author: "Caleigh Charlebois"
date: "2022-09-15"
output: pdf_document
latex_engine: "xelatex"
geometry: margin=1in
csl: "../references/biology-letters.csl"
bibliography: "../references/lizard_learning_plot.bib"
nocite: '@*'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Use groundhog to ensure package compatibility by loading most recent
## versions of packages available on a certain date.

# Check whether groundhog is installed and install it if it isn't.
my_packages <- c("groundhog")
not_installed <- my_packages[!(my_packages %in% installed.packages()[ , "Package"])]
if(length(not_installed)) install.packages(not_installed) 
rm(not_installed, my_packages)

# Load groundhog and use it to install other packages. You may change the date
# in the "date_for_packages" if you want to check results using another set of
# plugin versions.
date_for_packages <- "2022-09-11"
library(groundhog)
groundhog.library(c("ggplot2",
                    "readr",
                    "cowplot",
                    "dplyr"),
                  date_for_packages)
rm(date_for_packages, r.need.full)

# Read vector with names of data files (with dates added by data downloading 
# script) that can be indexed to get the names of both data files in the
# directory.

data_names <- read_lines("../data/data_names.txt")

```

**Title:** Lizard Learning Plot

Caleigh Charlebois


# Introduction
Most research on Australian skinks has focused on conservation strategies<citation> and their reproductive anatomy as live-bearing reptiles<cite>. In contrast, the data I am exploring and visualizing in this project records a phenomenon that is often dismissed in reptiles: social learning<citation>. Though reptiles are less likely than mammals and birds to live social lifestyles, this paper posits that there may still be a selective advantage to social learning: Lizards may be able to glean information about accessing the resources and avoiding the dangers of their environment by watching other individuals, even if they are otherwise antisocial<citation>.
In this study, researchers tested skinks' performance in two social learning tasks. In both tasks, lizards were placed in an enclosure with a food motivation (a mealworm) underneath a colored lid. The task was to flip the lid over to access the worm inside.
In the experimental group, the skinks were allowed to watch a trained demonstrator skink engage in the tasks for one hour prior to attempting the tasks themselves. The control group was allowed to observe the demonstrator lizard in its enclosure but not performing the tasks.
Task 1 was an instrumental task, in which a single yellow lid was present and needed to be overturned to access the worm. Task 2 was a more involved association task, in which lizards had to select either a blue or white lid, with only the blue lid hiding an accessible worm. 
The authors concluded that young lizards seemed to benefit from social learning by performing the task more quickly when they had the opportunity to observe the demonstrator lizard performing it first, while older lizards had similar outcomes whether they observed the demonstrator lizard performing the task or not.
In this project, I will be exploring the data researchers collected as part of task 1, also called the instrumental task. I will explore ways of visualizing the data using the ggplot2 package with specific attention to the facet_grid() function.

# Data exploration and Plots
```{r}
# Load the first lizard learning dataset

lizards_1 <- read_csv(paste0("../", data_names[1]))

ggplot( data = lizards_1, mapping = aes()) +
  geom_histogram(aes(x=LT)) +
  facet_grid(rows = vars(Age))
```
This was my initial plan -- Making a plot of the distribution of number of trials it took for the lizards to perform the task correctly, faceted by age. However, without polishing the plot and fixing the labels, I can see it won't be very exciting and isn't the best way to visualize this data.I didn't realize when I was making my plan that the  lizards would all learn the task in either 0 or 1 trial. 

What else can I make a plot of which will be more interesting?

```{r}
# Create a labelling function that will allow me to choose more readable facet
# labels for  "Age" and "Treatment" functions. Otherwise, the facets will be
# labeled using the abbreviations the authors chose. For example, simply "O"
# for "Old" and "SL" for "social demonstration."

age_labels <- list(
  'Y'='Young',
  'O'='Old'
)

plot_labeller <- function(variable, value){
  if (variable=='Age') {
    return(age_labels[value])
  } else {
    return(treat_labels[value])
  }
}

treat_labels <- list(
  'SL' = 'Social Demonstration',
  'C' = 'Control'
)

treat_labeller <- function(variable, value){
  return(treat_labels[value])
}

# Create a plot of the data with ggplot, facet with the facet_grid function,
# and use the cowplot theme. Specify the labeler I created above.

ggplot(data = lizards_1, mapping = aes()) +
  geom_histogram(aes(x=Lat_min1)) +
  facet_grid(rows = vars(Age),
             cols = vars(Treat),
             labeller = plot_labeller) +
  labs(x="Time until correctly performed task",
       y="Lizard count") +
  theme_cowplot()

```
I chose to make the x axis the time elapsed until the lizard correctly performed the task rather than simply whether they performed it correctly or not. I also changed the theme of the plot to the cowplot theme, which is visually cleaner.

I find this idea pretty interesting, but I notice there are many data points at "60." I think this is because lizards who never completed the task correctly were recorded as taking 60 seconds, the maximum allotted time. Next, I will try creating a plot that removes lizards which never succeeded in performing the task by filtering the data.

```{r}
filter(lizards_1, correct.incorrect==1) %>%

ggplot(mapping = aes()) +
  geom_histogram(aes(x=Lat_min1),
                 binwidth = 1) +
  facet_grid(rows = vars(Age),
             cols = vars(Treat),
             labeller = plot_labeller) +
  labs(x="Time until correctly performed task",
       y="Lizard count") +
  theme_cowplot()
```

This plot gives a better idea of the distribution of task completion times for young and old lizards. Based on this plot alone, it appears that a higher percentage of old lizards performed the task more quickly when social demonstration was available, while younger lizards tended to perform the task more quickly without social demonstration. However, it is worth noting that in this step, I removed lizards that didn't learn the task at all.

```{r}
filter(lizards_1, Liz.ID=="*0302")
```
Above, I have filtered the data set by the "Liz.ID" variable because I was curious about the structure of the data. This has revealed something I didn't understand about the data before: Each entry is not a lizard, it's a specific trial the lizard underwent. Even though the data dictionary available on Dryad says that the "LT" or "Learning Trial" variable is "trials until the lizard learnt," it is only a binary entry. 
Maybe this means the variable is "1" when the lizard hasn't yet learned the task, and "0" once it has learned the task? I still find this confusing because the example lizard here appears to have correctly performed the task during trial 4 according to the "correct.incorrect" column. 
Because I am not sure how to determine on which trial a lizard learned the task, I no longer want to work with the "learning trial" variable. Instead, what if I explore whether the demonstrator used to teach a lizard the task appears to impact the time it takes for the lizard to perform the task itself?

```{r}

lizards_1$correct.incorrect <- as.factor(lizards_1$correct.incorrect)

ggplot(data = lizards_1, mapping = aes(x=Dem.ID, y=Lat_min1)) +
# Create a jitter plot where points are colored by whether they represent
# successful trials
  geom_jitter(position=position_jitter(0.1),
              aes(col=correct.incorrect),
              alpha=0.5) +
# Define the color scale and the labels it will have in the legend
  scale_color_manual(values = c("red", "black"),
                     labels = c("Incorrect", "Correct")) +
# Use the cowplot theme for a cleaner result
  theme_cowplot() +
# Change the labels for the plot and add a title
  labs(title="Task duration grouped by demonstrator",
       x="Demonstrator ID",
       y="Seconds until task performed",
       col="Task result") +
# Rotate the x axis labels by 90 degrees so that they do not overlap
  theme(axis.text.x=element_text(angle= 90,
                                 vjust= 0.5,
                                 hjust= 1))

ggplot(data = lizards_1, mapping = aes(x=Dem.ID, y=Lat_min1)) +
# Create a boxplot where points are colored by whether they represent
# successful trials
  geom_boxplot() +
# Use the cowplot theme for a cleaner result
  theme_cowplot() +
# Change the labels for the plot and add a title
  labs(title="Task duration grouped by demonstrator",
       x="Demonstrator ID",
       y="Seconds until task performed",
       col="Task result") +
# Rotate the x axis labels by 90 degrees so that they do not overlap
  theme(axis.text.x=element_text(angle= 90,
                                 vjust= 0.5,
                                 hjust= 1))
```
I like how these two plots came out! I decided to plot the data both as a jitter plot and a box plot, because the box plot made it difficult to gauge how many failed attempts were clustered together at the top of the plot. The box plot makes it clearer that the lizards following each demonstrator didn't perform the same. Some of the demonstrators, specifically 198174 and 201191, seemed to have a lot of success in demonstrating the task because many of their followers did it quickly. In contrast, others had most of their followers not succeed at all.

```{r}

# Use the filter function to remove trials in which lizards were unsuccessful
filter(lizards_1, correct.incorrect==1) %>%

ggplot(mapping = aes(x=Dem.ID, y=Lat_min1)) +
  geom_boxplot() +
  theme_cowplot() +
  labs(title="Task duration grouped by demonstrator",
       x="Demonstrator ID",
       y="Seconds until task performed",
       col="Task result") +
  theme(axis.text.x=element_text(angle= 90,
                                 vjust= 0.5,
                                 hjust= 1))
```
When unsuccessful trials are removed, as above, there is still quite a bit of variation. I am surprised by the distribution of times here: It seems like, for the most part, lizards performed the task quickly if they were going to do it correctly rather than forming a normal distribution around an intermediate average time. 

# Discussion
Though this exercise reinforced my understanding of the ggplot plotting syntax and dplyr functions, my main takeaway was actually the difficulty I had in interpreting the meaning of the variables in the tabular data. When I chose this dataset, I thought it would be easy to understand because I saw that the authors had taken the time to describe their variables on dryad. When I actually went to work with the data, however, I found that some variables were still not adequately explained. Most problematically, I wasn't able  to figure out how the "LT" and "correct.incorrect" columns combined to characterize how many trials it took before a lizard learned the task. Not understanding the variables makes it difficult to know which apparent discrepancies are due to errors in the data and which ones are legitimate unexpected results.

The fact that the authors reported the duration of failed tasks as 60 seconds also made the data more difficult to work with. I would expect failed tasks to have a time of "N/A" based on their description of the variable, which was "time from start of trial to the point when the lizard successfully flipped the yellow lid."

I would also like to acknowledge the discrepancy between the plot that I am happy with and what I anticipated I would be doing when I created my preregistration. In the preregistration, I predicted that "lizard age [would] positively correlate with performing the lid-flipping task correctly on the first attempt." In my final manuscript, I have instead visualized how the distribution of times it has taken for the lizards to perform their tasks varies when the task was demonstrated with different demonstrator lizards. In a typical study, a departure from the pre-registration based on which data seemed most interesting to plot would be ridiculous. However, I feel comfortable departing in this case because the true objective of the project is to become comfortable with reproducible science.

Overall, I consider this project a success. I have become more comfortable with skills that I think will make my thesis project better and more reproducible. For example, I have been creating pipeline scripts to use bioinformatics tools on the Compute Canada cluster, and currently have no record of the ways I have edited these scripts over time. When I accidentally break something or need to debug a script, I am in a precarious position of needing to remember what I did before. This course has reminded me how valuable it will be to create a git repository tracking changes in the scripts I create on the cluster, and properly document the purposes and provenances of my scripts with readme files. 

Additionally, I previously saw R Markdown documents as relatively rudimentary, and hadn't seen the visual editor in RStudio at all. Through conducting this project, I learned how my workflow could integrate my citation manager with R Markdown citations. I also had the opportunity to see that R Markdown manuscripts are more customizable than I previously thought.

# References
